---
title: "COVID-19 State Community Mobility Map"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  html_document:
    code_folding: hide
  github_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(maps)
library(mapdata)
library(ggmap)
library(plotly)
library(patchwork)
```

## Map and data setup
Data table of percent change visits extracted from [Google Community Mobility Report](https://www.google.com/covid19/mobility/), based on data from **04/05/2020**.

*Expand code to see data and map setup*
```{r data}
# ===  Loading data from "maps" and community mobility data report === #

# * Loading map from `library(maps)` for drawing the maps.
# * Load state's community mobility data from google's PDFs (extracted with `scripts/getdata.R`).
#   + Note: There were data available for Alaska and Hawaii, but since it won't be mapped, they were removed here.
  
# 1. load maps
usa <- map_data("usa")
states <- map_data("state")

theme_map <- function(...) {
  theme_minimal() +
  theme(
    axis.line = element_blank(),
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank(),
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    # panel.grid.major = element_line(color = "#ebebe5", size = 0.2),
    panel.grid.major = element_blank(),    
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    ...
  )
}

# 2. Load dataframe(s) into a list
file_list <- paste0("./data/", dir("./data")[grep(".csv", dir("./data"))])

commlist <- list()
for(i in 1:length(file_list)){
  commlist[[i]] <- read.csv(file_list[i])
  
  # take out alaska & hawaii just for mapping # also we don't have DC
  commlist[[i]] <- commlist[[i]] %>% filter(!is.element(commlist[[i]]$states, c("Alaska", "Hawaii"))) 
  
  # Rename states to in community mobility data to match map
  commlist[[i]]$states <- tolower(commlist[[i]]$states)
  commlist[[i]]$states <- gsub("_", " ", commlist[[i]]$states)
  
  fname <- sub(pattern = "(.*)\\..*$", replacement = "\\1", basename(file_list[i])) %>%  # Get date of first data file
                        strsplit(x = ., "_")
  names(commlist)[i] <- format(as.Date(fname[[1]][5], format = "%Y%m%d"), "%Y-%m-%d")
}

# bdf <- read.csv(file_list[1]) # load first data file
# 
# fname <- sub(pattern = "(.*)\\..*$", replacement = "\\1", basename(file_list[1])) %>%  # Get date of first data file
#   strsplit(x = ., "_")
# bdf$date <- format(as.Date(fname[[1]][5], format = "%Y%m%d"), "%Y-%m-%d")
# 
# 
# for(i in 2:length(file_list)){ # combine data from multiple dates into one dataframe
#   odf <- read.csv(file_list[i])
#   fname <- sub(pattern = "(.*)\\..*$", replacement = "\\1", basename(file_list[i])) %>% 
#     strsplit(x = ., "_")
#   
#   odf$date <- format(as.Date(fname[[1]][5], format = "%Y%m%d"), "%Y-%m-%d")
#   
#   bdf <- rbind(bdf, odf)
# }


# ===  Wrangled data to form data frame for plotting === #
# * Merge data from community mobility report to state's lat/long data frame
# * Reshape data from wide to long format for plotting
# * Rename/recode variables

ndata <- length(commlist)

datlist <- list()

for(nd in 1:ndata){
  ns <- data.frame(states, retail=NA, grocery=NA, parks=NA, transit=NA, workplace=NA, residence=NA)
  
  
  for(i in 1:nrow(ns)){
    if(ns$region[i]=="district of columbia"){
        ns[i,7:ncol(ns)] <- NA
    }else{
        ns[i,7:ncol(ns)] <- as.character(unlist(commlist[[nd]][commlist[[nd]]$states==ns$region[i],2:ncol(commlist[[nd]])]))
    }
  }
  
  for(j in 7:ncol(ns)){ 
    ns[,j] <- as.numeric(sub("%", "",ns[,j])) # remove percentage sign
  }

  ns[,7:ncol(ns)] <- ns[,7:ncol(ns)] *.01 # convert whole number to *.01 for plotting as %
  
  ns$date <- names(commlist)[nd]
  
  #  Gather to long form
  ldf <- ns %>% 
    gather(data = ., key = "Type", value = "Percent_Change", retail:residence, factor_key = T)
  
  ldf$Type<- recode(ldf$Type,
           retail="Retail & recreation", 
           grocery="Grocery & pharmacy",
           parks="Parks", 
           transit="Transit stations", 
           workplace="Workplaces", 
           residence="Residential")
  
  datlist[[nd]] <- ldf
  names(datlist)[nd] <-  names(commlist)[nd]
}

```

# Maps

**The colors in the map indicates % change in visits to  different categories of places (e.g., grocery, parks), comparing Mar.29 to a baseline value = median of Jan 3-Feb 6**

## 1. Reduced visits (community mobility) compared to Jan3-Feb6 in these places: 

* Retail & recreation
* Grocery & pharmacy
* Transit stations
* Workplaces

### % change in visits to Retail/Grocery/Transit/Workplaces
```{r reduced_map, dev='svg', message=F}
maplist <- list()

for(nd in 1:ndata){
  
 lp <- datlist[[nd]] %>% 
    filter(!is.element(Type, c("Parks", "Residential"))) %>%
    ggplot(aes(text = region)) + 
      geom_polygon(aes(x = long, y = lat, group = group, fill=Percent_Change), color = "grey80") + 
      coord_fixed(1.3) +
      scale_fill_gradient2(low = "navyblue", mid = "white", high="firebrick", midpoint = 0, lim=c(-.8,.1),
                           labels = scales::percent) +
      labs(fill = "% change in \nvisits") +
      facet_wrap(~Type, nrow =1) +
      # annotate(unique(datlist[[nd]]$date)) +
      theme_map()
 
  maplist[[nd]] <- ggplotly(lp) # make text appear on click/hover
}


# ggsave(filename = "./figs/reduced_map.png", lp) # for README

# lpy <- ggplotly(lp)
# 
# lpy

subplot(maplist, nrows = ndata, titleX = F, titleY = F)


```

## 2. Increased visits (community mobility) in Residential areas

* All states showed increased density in residential compared to Jan3-Feb6.

### % change in visits to Residential areas

* 'visits' is probably a weird term here. 

```{r increased_map, dev='svg', message=F}
lp <- ldf %>% 
  filter(is.element(Type, c("Residential"))) %>%
  ggplot(aes(text = region)) + 
    geom_polygon(aes(x = long, y = lat, group = group, fill=Percent_Change), color = "grey80") + 
    coord_fixed(1.3) +
    scale_fill_gradient(low = "white",high = "firebrick", lim=c(0,.25),
                         labels = scales::percent) +
    # facet_wrap(~Type) + 
    labs(fill = "% change in \nvisits") +
    ggtitle("\n\nResidential") +
    theme_map() +
    theme(plot.title = element_text(hjust = 0.5))


ggsave(filename = "./figs/increased_map.png", lp) # for README

lpy <- ggplotly(lp)

lpy
```

## 3. Mixed increase/decrease in vistis in Parks 

* States such as Kansas and North Dakota showed huge increases in visits to Parks compared to the median baseline (Jan 3-Feb 6). 
* Other states such as California and Florida showed decreases. 

### % change in visits to Parks
```{r mixed_map, dev='svg',message=F}
lp <- ldf %>% 
  filter(is.element(Type, c("Parks"))) %>%
  ggplot(aes(text = region)) + 
    geom_polygon(aes(x = long, y = lat, group = group, fill=Percent_Change), color = "grey80") + 
    coord_fixed(1.3) +
    scale_fill_gradient2(low = "navyblue", mid="white",high = "firebrick", lim=c(-.8,1.),
                         labels = scales::percent) +
    # facet_wrap(~Type) + 
    labs(fill = "% change in \nvisits") +
    ggtitle("\n\nParks") +
    theme_map() +
    theme(plot.title = element_text(hjust = 0.5))

ggsave(filename = "./figs/mixed_map.png", lp)

lpy <- ggplotly(lp)

lpy
```



# Data

* Data table of actual values plotted. These values were pulled from https://www.google.com/covid19/mobility/ based on data from 3/29/2020.

```{r datatable}
df %>%
  DT::datatable( caption = "Google Community Mobility Data - 3/29/2020") 
```

